### 2018-03-12 docs/paper.tex
more on docs

%%% template.tex
%%%
%%% This LaTeX source document can be used as the basis for your technical
%%% paper or abstract. Intentionally stripped of annotation, the parameters
%%% and commands should be adjusted for your particular paper - title,
%%% author, article DOI, etc.
%%% The accompanying ``template.annotated.tex'' provides copious annotation
%%% for the commands and parameters found in the source document. (The code
%%% is identical in ``template.tex'' and ``template.annotated.tex.'')

\documentclass[]{acmsiggraph}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\TOGonlineid{45678}
\TOGvolume{0}
\TOGnumber{0}
\TOGarticleDOI{0}
\TOGprojectURL{}
\TOGvideoURL{}
\TOGdataURL{}
\TOGcodeURL{}
\usepackage{color}
%\definecolor{red}{rgb}{0.9, 0.17, 0.31}
\usepackage{multirow}
\usepackage{subfig}
\usepackage{xcolor}
\usepackage{lipsum}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{glsllst} % My own package providing markup listing for glsl
\usepackage{rmlst}   % My own package providing markup listing for renderman
\usepackage{amsmath}
\usepackage{hyperref}

\lstset{
	backgroundcolor=\color[rgb]{0.95, 0.95, 0.95},
	tabsize=3,
	%rulecolor=,
	basicstyle=\footnotesize\ttfamily,
	upquote=true,
	aboveskip={1.5\baselineskip},
	columns=fixed,
	showstringspaces=false,
	extendedchars=true,
	breaklines=true,
	prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
	frame=none,
	aboveskip=15pt,
	belowskip=8pt,
	captionpos=t,
	showtabs=false,
	showspaces=false,
	showstringspaces=false,
	identifierstyle=\ttfamily,
	%keywordstyle=\color{red}\bfseries,
	%keywordstyle=[1]\bfseries\color{syntaxBlue},
	%keywordstyle=[2]\bfseries\color{syntaxRed},
	%keywordstyle=[3]\color{blue}\bfseries,
	%keywordstyle=[4]\bfseries\color{syntaxBlue},
	commentstyle=\color[rgb]{0.082,0.639,0.082},
	keywordstyle=[1]\bfseries\color[rgb]{0,0,0.75},
	keywordstyle=[2]\bfseries\color[rgb]{0.5,0.0,0.0},
	keywordstyle=[3]\bfseries\color[rgb]{0.127,0.427,0.514},
	keywordstyle=[4]\bfseries\color[rgb]{0.4,0.4,0.4},
	stringstyle=\color[rgb]{0.639,0.082,0.082},
}

\title{Innovations Report: Procedural Fur System}

\author{Joe Withers\thanks{e-mail:joewithers96@gmail.com}\\National Centre for Computer Animation}
\pdfauthor{Joe Withers}

\keywords{rendering}

\begin{document}

\maketitle

\begin{abstract}
For this project I developed a procedural fur system, with the aim of exploring the feasibility of offloading computation onto the GPU within artist tools. The final artefact is an application which serves to interface with the fur system API that I have developed. This report primarily documents the implementation of the API, as well as my findings.
\end{abstract}

\section{Introduction} \label{sec:introduction}
Introduction Text.

\subsection{Existing Solutions} \label{sec:existing}
Whilst many larger companies opt to develop their own software for handling fur, such as MPC's Furtility \cite{furtility}, there are a number of commercially available solutions that are also widely used in production. The most popular of these is XGen, originally developed by Walt Disney Animation Studios, which has been included in Autodesk's Maya since 2014. XGen uses a 'Collection' and 'Description' paradigm; Descriptions store user controlled parameters which affect fur primitives (curves for example) on a select area of a mesh, and Collections act as a container for these Descriptions. There are two primary methods offered for distributing primitives on a mesh, Guide Curves can be used to manually control the placement and shape of primitives by placing 'Guides' on the mesh surface, whereas 'Groomable Splines' randomly distirbutes splines over the mesh surface which can be manipulated in real time using viewport manipulators that mimic the brushing and styling of fur. In both modes, 'modifiers' can be added to the modifier stack to further adjust the visual outcome, adding features such as clumping, noise, and coiling. Attributes within these modifiers can be further controlled using expressions or texture maps.

Another commercially available solution is Peregrine*Labs' Yeti \cite{yeti}, which is also offered as a plugin within Autodesk's Maya. Yeti offers a more procedural approach, giving the user a node graph interface in which to construct the desired look from individual components. For example, the most simple usage requires the user to first create a 'scatter' node to distribute points on the mesh surface, and then combine it with a 'grow' node to extrude curves from these points. Whilst at first this may seem unintuitive when compared to XGen, it offers greater flexibility and avoids the pitfalls associated with a destructive workflow.

Offerering even even more flexibility, recent versions of SideFX's Houdini include hair and fur operator nodes, which create predefined node networks for the generation of hair and fur. These networks use standard Houdini nodes internally, so experienced Houdini users are intuitively able to create the visual outcome they desire, using tools they already know. However, it may be quite unintuitive for asset artists (those who would be using the tool, Groom artists?), who are not used to Houdini's heavily procedural workflow.

\section{Implementation} \label{sec:implementation}
Method Text.

\subsection{Resources} \label{sec:resources}
I decided to develop the API side of the fur system using C++, primarly as it most commonly used when developing computationally heavy artist tools, but also because it is the language I am most comfortable using. I opted to use OpenGL over other APIs (CUDA, OpenCL) for offloading of computation onto the GPU, simply because my final artefact is a tool with a graphical interface, and OpenGL is capable of handling both arbitary computation using compute shaders and rendering of geometry simultaneously.

To handle the user interface I used the Qt framework within C++. Qt is commonly used for artist tools within visual effects as it is cross-platform, and applications can be configured to run within other applications that make use of it, such as Autodesk Maya.

I wanted to include a node-graph style interface within my application, as it is commonly used within existing artist tools (Autodesk Maya, Unreal Engine), and would encourage modularity within my API. Qt does not natively provide this kind of interface, so I made use of NodeEditor \cite{Pinaev2017}, an existing Qt-based library that provides this functionality.

\subsection{Design} \label{sec:design}
From looking at existing fur systems I was able determine that my simplified fur system would need to consist of the following components:

\begin{itemize}
	\item Geometry Loaders - These are responsible for loading user specified geometry. Example Geometry Loaders could allow for the parsing of Wavefront OBJ files.
	\item Distributors - These are responsible for the distribution of curves onto user specified geometry. User controllable parameters could include density (curve count), distribution pattern (random, uniform), and curve length. These parameters could potentially be controlled by texture inputs.
	\item Operators - These are responsible for manipulation of the curves to achieve the desired look. Example operators could provide bending, clumping, or randomisation of input curves.
	\item Renderers - These are responsible for the rendering of curves into the application viewport. Example renderers could provide mesh, cuves as lines, or curves as ribbons rendering functionality. User controllable parameters could provide controls for the shading model in use, as well as control of the base and tip widths when rendering curve ribbons for example.
\end{itemize}

Whilst the list of components specified above provide the required functionality for a standalone tool, they do not provide any functionality for exporting fur for use in a broader pipeline. I chose to omit this functionality to limit the scope of this assignment, but the ability to export hair curves as alembic would be desirable, as alembic is a widely adopted format that allows for the storage of multiple different types of curve geometry. Another candidate file type would be RenderMan's RIB format, which would allow for direct rendering of the hair curves in any RenderMan compliant renderer.

\subsection{Development} \label{sec:development}

write about obj loading

write about distributing https://stackoverflow.com/questions/9294316/distribute-points-on-mesh-according-to-density

write about GPU

\subsection{User Testing} \label{sec:usertesting}
Near the end of the development I was able to convince friend and fellow student Alin Bolcas to test the system and provide feedback from an artist's point of view; Alin is an experienced modelling and look development artist who uses a variety of different software packages to create his work, so getting his feedback on the usability of my tool proved extremely useful.

His first comments were in regards to how the node graph handles rendering of meshes and curves; he found it quite unintuitive to manually create rendering nodes for each new curves object that needed to be drawn. He suggested that a node for the combination of curves should be added, which would consolidate multiple curve objects into a single output for simplified rendering.

He noted that the lack of a node that applies noise to curves proved problematic during his testing; this kind of node is fundamental to achieving appealing fur visuals. He liked the behaviour of the clumping node, as this functionality is also very common and highly used within fur systems, though visual quality was somewhat compromised as my current implementation does not preserve the initial curve length. Adding a control in the distributor that would allow for the variation of curve length was also suggested as something that would improve the quality of the visual output.

His final comments were in regards to the user interface, suggesting that I make controls similar to their counterparts in existing software wherever possible. Changing the camera controls to match those in Maya was suggested as it would provide a familiar user experience when manipulating the viewport. The ability to create new nodes by pressing the tab key as opposed to right click was also suggested, as this is the standard in node editors, such as those in Maya and Nuke.

\section{Results} \label{sec:results}
Results Text.

\subsection{Performance} \label{sec:performance}

insert stats comparing CPU/GPU compute

\section{Conclusion} \label{sec:conclusion}

\bibliographystyle{acmsiggraph}
\bibliography{references}

\end{document}